---
title: "IBD_Analysis"
author: "David Hu"
date: "2022-11-24"
output:
  word_document: default
  pdf_document:
    latex_engine: xelatex
header-includes:
  \usepackage{fvextra}
  \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

# IBD Analysis 

```{r}
#load in data
data = read.csv("~/Documents/nsqip_ibd_ic/data/comb5_data.csv",header = TRUE)
data <- replace(data, data == "", NA)
#data$ileostomy.formation  <- ifelse(test = data$ileostomy.formation =="Ileostomy", yes = 1, no = 0)
```
Replaced all missing values with "NA" for imputing values later. 

```{r}
#convert all character variables to factor variables
data[sapply(data,is.character)]<-lapply(data[sapply(data,is.character)],as.factor)
#various methods to check for possible columns to drop
#sapply(lapply(data, unique), length) #display number of levels
#sapply(data[, sapply(data, class) == 'factor'], levels) #levels of each variables
#mapply(table, data) #display levels and count for each level

```
Convert the character variables to factor variables to represent categorical data. 

```{r}
#colSums(data[, sapply(data, class) == 'factor']=="")/nrow(data) old method
#sapply(data[, sapply(data, class) == 'factor'], levels)
#colSums(is.na(data))/nrow(data)
```
The "sapply(data[, sapply(data, class) == 'factor'], levels)" line is used to display the amount of levels for a categorical variable. We can find which what variables have one level so we can drop them because they will not be useful when building our model. 

The "colSums(is.na(data))/nrow(data)" displays the proportion of missing values for each dataset. If a categorical variable has 25% or more of its values missing, then we drop the variable. 

```{r}
#from Kevin's email
data = subset(data,select=-c(X..of.postop.superficial.incisional.ssi,X..of.postop.deep.incisional.ssi,X..of.postop.organ.space.ssi,X..of.postop.wound.disruption,X..of.postop.pneumonia,X..of.postop.unplanned.intubation,X..of.postop.pulmonary.embolism,X..of.postop.on.ventilator...48.hours,X..of.postop.progressive.renal.insufficiency,X..of.postop.acute.renal.failure,X..of.postop.cva,X..of.postop.uti,X..of.postop.cardiac.arrest.requiring.cpr,X..of.postop.myocardial.infarction,X..of.postop.transfusion.intraop..postop..72h.of.surgery.start.time.,X..of.postop.vein.thrombosis.requiring.therapy,X..of.postop.c..diff,X..of.postop.sepsis,X..of.postop.septic.shock,X..of.postop.other.occurrences,surgical.wound.s..closure,hospital.length.of.stay,hospital.discharge.destination,postop.death.w.in.30.days.of.procedure,postoperative.death...30.days.of.procedure,X..of.readmissions.w.in.30.days,X..of.unplanned.readmissions,first.unplanned.return.cpt,first.unplanned.return.to.or,colectomy.postop.anastomotic.leak,colectomy.prolonged.postoperative.npo.or.ngt.use))

#drop variables that have >=25% missing values and those with only one levels
data = subset(data, select = -c(malnourishment,colectomy.malignancy..pathologic.m.stage,colectomy.malignancy..pathologic.n.stage,colectomy.malignancy..pathologic.t.stage,
preop.transfusions..rbc.w.in.72.hrs.prior.to.surgery.start.time.,
congestive.heart.failure.w.in.30.days.prior.to.surgery,ascites.w.in.30.days.prior.to.surgery,
colectomy.primary.indication.for.surgery,
colectomy.non.emergent.indication.icd10.code,
currently.requiring.or.on.dialysis,
acute.renal.failure,
case.number
))
```

We've identified the following variables to drop based on whether <=25% of its values were missing and/or had only one level. While case.number didn't fit the criteria, it was dropped because it doesn't provide relevant information since it is the surgery case number.  

```{r, warning=FALSE}
#imputing the missing values
#install.packages("missForest")
library(missForest)
data.imp <- missForest(data)$ximp
```

Impute missing values using random forest nonparametric modeling.

```{r,warning=FALSE}
#stepwise logistic regression 
library(MASS)
model <- glm(ileostomy.formation~.,data=data.imp,family="binomial")
step.model <- stepAIC(model,trace = FALSE,direction = "backward")
summary(model)
```

Use step wise AIC logistic regression with backward selection to select the variables that are the most useful to help the explain the probability of an illeostomy after logistic regression (will be referred to as original logistic model) on all the variables throws a warning. 

```{r}
library(car)
#vif(model)
```
Unable to check multicollinearity using vif (Variance Inflation Factor) on the original logistic model. 
Throws warning "Error in vif.default(model) : there are aliased coefficients in the model".

```{r}
probabilities <- predict(model, data.imp, type = "response")
predicted.classes <- ifelse(probabilities > 0.50, "None","Ileostomy")
# Prediction accuracy
observed.classes <- data.imp$ileostomy.formation
mean(predicted.classes == observed.classes)
```

Although original logistic model throws a warning, the model does predict the correct outcome for 87% of the observations. 

```{r}
summary(step.model)
```
Coefficients for the stepwise AIC logistic model. 

```{r}
vif(step.model)
```
Using VIF, no multicolinearity detected (VIF>=5)

```{r}
probabilities <- predict(step.model, data.imp, type = "response")
predicted.classes <- ifelse(probabilities > 0.5, "None", "Ileostomy")
# Prediction accuracy
observed.classes <- data.imp$ileostomy.formation
mean(predicted.classes == observed.classes)
```
The accuracy does go down by about one percent using the step wise AIC logistic model. 

```{r}
#the coefficients for each model ran in here
as.data.frame(coef(model))
as.data.frame(coef(step.model))
```

```{r}
#other stuff to experiment with it
#levels(data.imp$race) <- c("Other","Other","Black or African American","Other","Other","Unknown/Not Reported","Other")
#levels(data.imp$origin.status) <- c("Other", "From acute care hospital inpatient", "Other", "Not transferred (admitted from home)","Other","Other","Other","Other")
#levels(data.imp$wound.classification) <- c("Other","Other","Other","Contaminated","Other")
```

Bottom was my attempt at building a model if we wanted to some sort of modeling for future predictions but not needed for this context. Tried to combine levels to make all the variables in model significant. 

